An integrated open population distance sampling approach for modelling age-structured populations

### 2024-01-05 Manuscript_Nilsen_Nater.qmd
Updated acknowledgments, data/code availability, added code ref

Compared to previous studies on ptarmigan (see e.g. [@Israelsen2020; @Sandercock2011]), we could have expected somewhat higher estimates of survival probability. One potential reason for the overall lower estimates obtained here is that our IDSM analysis assumed constant survival over the period 2007-2021 while survival, in reality, may have changed over time. If survival in more recent years, when telemetry data was collected [and the study of e.g. @@Israelsen2020 was carried out] was higher than in earlier years -- something that seems likely given population increase over recent years -- an average over the entire time period is expected to be lower. On a different note, we can also not exclude the possibility of a small degree of bias in survival estimates due to misclassification of age in the data (see above), especially seeing as the IDSM's recruitment rate estimates also appear somewhat high compared to other studies on ptarmigan (+refs). <!-- If we have any "biological" thoughts about why somewhat high recruitment rates may not be that ridiculous here, then that could be added here -->

@@

>>>>

Compared to previous studies on ptarmigan (see e.g. [@Israelsen2020; @Sandercock2011]), we could have expected somewhat higher estimates of survival probability. One potential reason for the overall lower estimates obtained here is that our IDSM analysis assumed constant survival over the period 2007-2021 while survival, in reality, may have changed over time. If survival in more recent years, when telemetry data was collected [and the study of e.g. \@ @Israelsen2020 was carried out] was higher than in earlier years -- something that seems likely given population increase over recent years -- an average over the entire time period is expected to be lower. On a different note, we can also not exclude the possibility of a small degree of bias in survival estimates due to misclassification of age in the data (see above), especially seeing as the IDSM's recruitment rate estimates also appear somewhat high compared to other studies on ptarmigan (+refs). <!-- If we have any "biological" thoughts about why somewhat high recruitment rates may not be that ridiculous here, then that could be added here -->

@@

The raw data from the line transect surveys is deposited on GBIF and can be accessed freely via the Living Norway Data Portal (https://data.livingnorway.no/). The work here is based on version 1.7 of the dataset "Tetraonid line transect surveys from Norway: Data from Fjellstyrene" (key b49a2978-0e30-4748-a99f-9301d17ae119).

The auxiliary radio-telemetry data and rodent occupancy data, and all code used for wrangling, analysing, and visualizing data and results can be found in the project's repository on GitHub: <https://github.com/ErlendNilsen/OpenPop_Integrated_DistSamp>. The results presented in this paper were created using version 1.3 of the code [@IDSMCode_2024].

Supplementary figures and posterior samples from the model run on the real data are archived on OSF: (**TBA**)<!-- Add link/reference to OSF storage -->
